---
title: "Agroecological Dynamics"
output: 
  html_document: 
    highlight: haddock
    keep_md: yes
---
Agent-based model of Roman land use and agroecology in North Africa
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```
# Setup
Import necessary packages
```{r libraries, message = F, warning = F}
library(raster)  # for raster manipulation
library(MODIS) # land cover data
library(tidyverse) # for data manipulation and plotting
library(mgcv) # for fitting functions to data
library(viridis) # for color palettes
library(gridExtra) # for arranging plots
library(pbapply) # for progress bar and parallelization
library(pleiades)
library(gdistance)
```

# Modules
First we define key functions dealing with farming, soil dynamics, and demography.

### Soil Dynamics

Define parameters controlling the soil fertility dynamics.
```{r soil_parameters, eval=FALSE, include=FALSE}
carrying.capacity <- 100
degradation <- tibble(type = c('reversible', 'hysteresis', 'irreversible'), 
                      degradation_factor = c(0, 1, 2),
                      regeneration_rate = c(0.05, 0.0844, 0.11880))
depletion.rate <- 0.25 # .25  to represent biennial fallow, .5 otherwise
```

Soil fertility dynamics
```{r soil-dynamics, eval=FALSE, include=FALSE}
soil.dynamics <- function(x, population, soil.type = 'reversible'){
  params <- filter(degradation, type == soil.type)
  params$regeneration_rate * x * (x / carrying.capacity) ^ params$degradation_factor * (1 - x / carrying.capacity) - depletion.rate 
}
```

So if you have 3 or more households per sq km (or 6 or more households practicing biennial fallow) then the soil will always deplete.
```{r eval=FALSE, include=FALSE}
expand.grid(fertility = 0:100, population = 0:5, soil.type = degradation$type, stringsAsFactors = F) %>%
  mutate(rate = pmap_dbl(list(fertility, population, soil.type), soil.dynamics)) %>%
  mutate(soil.type = factor(soil.type, levels = c('reversible', 'hysteresis', 'irreversible'))) %>%
  ggplot(aes(fertility, rate, group = as.factor(population))) +
  facet_wrap(~soil.type) +
  geom_line(aes(color = population)) +
    geom_hline(yintercept = 0, linetype = 2) +
  labs(title = 'Impact of population density on soil fertility change', x = 'Soil fertility (%)', y = 'Rate of fertility change') +
  theme_minimal()
```
### Time Allocation
```{r}

allocate_time <- function(p = 1, b = 1, lf, j = 0.3, q = 0, r = 1, k = 0.4, a = 1, w = 0.2, le = 1 - lf){
  p * b * lf ^ j * (q + r) ^ k * a ^ (1 - j - k) + w * le
}

expand.grid(lf = seq(0,1,.01), q = seq(0,1,.01), r = c(.1, .5, 1)) %>%
  mutate(utility = pmap_dbl(list(lf = lf, q = q, r = r), allocate_time)) %>%
  ggplot(aes(lf, utility, group = q, color  = q)) +
  geom_line() +
  facet_wrap(~as.factor(r)) +
  theme_minimal()
```

```{r}
p = 1
b = 1
j = 0.3
q = 0
r = 1
k = 0.4
a = 1
w = 0.2
l <- 1
psi <- .2
e <- .125
qmax <- 1
#region 1
(j * p * b * a ^ (1-j-k) * r ^ k / w) ^ (1/(1-j))

#region 2
1 / (j + k) * (k * l + j * (psi - e) - 2 * j * e * r / qmax)

#region 3
psi - e

######
allocate_time <- function(p = 1, b = 1, j = 0.3, r = 1, k = 0.4, a = 1){

}
```

```{r}
maintain_infra <- function(lm, psi = .2, epsilon = .18, imax = 1){
  ifelse(0 <= lm & lm < (psi - epsilon), 0,
  ifelse((psi - epsilon) <= lm & lm <= (psi + epsilon), imax / (2 * epsilon) * (lm - psi + epsilon), imax))
}
psi = .2; epsilon = .18
tibble(lf = seq(0, 1, .001), i = maintain_infra(lf)) %>%
  ggplot(aes(lf, i)) +
    geom_vline(xintercept = psi + epsilon, linetype = 2, color = 'grey') +
  geom_line() +
  labs(title = 'Impact of infrastructure maintainance on performance',
       x = 'Labor devoted to maintaining infrastructure', y = 'Infrastructure performance') +
  theme_minimal()
```


### Farming
Agents determine how much land they need, calculate the yields from their land, remember these yields, and harvest from the land.

First define some parameters relating to food production and consumption.
```{r food_params}
max.yield <- 1000 # maximum possible wheat yield, in kg/ha, should range from 1000 - 2000

calorie.req <- 2582 * 365  # annual individual calorie requirement, derived from daily requirement
wheat.calories <- 3320 # calories in a kg of wheat
wheat.cal.proportion <- 1 # percent of individual's food calories coming from wheat
wheat.req <- calorie.req / wheat.calories * wheat.cal.proportion # kg of wheat to feed a person for 1 year

sowing_rate <- 135  # kg of wheat to sow a hectare (range 108 - 135 from Roman agronomists)
seed_proportion <- 135 / max.yield # proportion of harvest to save as seed for next year's sowing

labor.per.hectare <- 40 # person days per hectare
max.ag.labor <- 280 # maximum days per year an individual can devote to farming
memory_length <- 10 #  number of years a household remembers
```

Crop yields are determined by rainfall and soil fertility.
```{r yields}
calc_climatic_yield <- function(precipitation){
  max.yield * pmax(0, 0.51 * log(precipitation) + 1.03)  # annual precipitation impact on yields
}

calc_yield_reduction <- function(fertility, climate_yield){
  climate_yield * pmax(0, 0.19 * log(fertility / 100) + 1)  # fertility impact on yields
} 
#todo, replace with gams!
```


```{r echo=F, fig.wdith = 8}
p1 <- tibble(fertility = 1:100, yield = calc_yield_reduction(fertility, 2000)) %>%
  ggplot(aes(fertility, yield)) +
  geom_line() +
  scale_y_continuous(limits = c(0,3000)) +
  labs(title = 'Soil fertility impact on wheat yield', subtitle = 'Assuming 1m annual precipitation',
       x = 'Soil Fertility (%)', y = 'Wheat Yield (kg/ha)') +
  theme_minimal()

p2 <- tibble(precipitation = seq(0, 2, .1), yield = calc_climatic_yield(precipitation)) %>%
  ggplot(aes(precipitation, yield)) +
  geom_line() +
  scale_y_continuous(limits = c(0,3000)) +
  labs(title = 'Precipitation impact on wheat yield', subtitle = 'Assuming 100% soil fertility',
       x = 'Annual Precipitation (m)', y = 'Wheat Yield (kg/ha)') +
  theme_minimal()

grid.arrange(p1, p2, ncol = 2)
```

How agents decide the amount of land they need to farm. The determination of how much land is needed is a function of the household's occupants and labor availability.
```{r}
calc_land_req <- function(n_inhabitants, yield, fallow = T){
  wheat.req * n_inhabitants * (1 + seed_proportion) / yield * ifelse(fallow, 2, 1)
}
```

```{r echo=F}
expand.grid(n_inhabitants = 1:20, Management = c('Fallow', 'No Fallow')) %>%
  mutate(land = if_else(Management == 'Fallow', calc_land_req(n_inhabitants, max.yield, fallow = T), 
                                                calc_land_req(n_inhabitants, max.yield, fallow = F))) %>%
ggplot(aes(n_inhabitants, land, color = Management)) +
  geom_line() + 
  labs(title = 'Household size and land requirements', subtitle = 'Assuming yield of 1,500 kg/ha', 
       x = 'Household Size (inhabitants)', y = 'Farm land requirement (ha)') +
  theme_minimal()
```

```{r land_req}
max_cultivable_land <- function(laborers, available_area, fallow = T, type = 'asymptote'){
  potential_area <- max.ag.labor * laborers * ifelse(fallow, 2, 1) / labor.per.hectare
  if(type == 'unlimited') return(potential_area)
  if(type == 'step') return(pmin(available_area, potential_area))
  if(type == 'asymptote') return(available_area * (1 - exp(-potential_area / available_area)))
}
```

```{r echo = F}
tibble(laborers = seq(0,30,.1), 
       potential = max_cultivable_land(laborers, 100, type = 'unlimited'),
       Step = max_cultivable_land(laborers, 100, type = 'step'), 
       Asymptote = max_cultivable_land(laborers, 100, type = 'asymptote')) %>%
  gather(Function, value, 3:4) %>% 
  ggplot(aes(potential, value, color = Function)) +
  geom_line() +
  labs(title = 'Space limited cultivation', subtitle = 'Adapted from Puleston and Tuljapurkar (2008)', x = 'Potential area cultivated by laborers (ha)', y = 'Cultivated area (ha)') +
  theme_minimal()
```



Here households calculate how much land they need, pull the crop yield from the environment, remember the yield, and determine their harvests by multiplying the yield by the amount of land they have (also removing some of the crop to save as seed for next year).

first, for dev purposes, make a fake environment raster.
```{r}
allocate_land <- function(households){
  households %>% 
   mutate(land = calc_land_req(n_inhabitants, peak_end(yield_memory)),
          max_land = max_cultivable_land(laborers, area * arable_proportion * 100, fallow = T, type = 'asymptote')) %>%
    mutate(land = pmin(land, max_land)) %>%
    select(-max_land)
}
```

```{r farm_funs}
farm <- function(households){
  households %>%
    mutate(yield = climatic_yield,
           yield_memory = yield, #map2(yield_memory, yield, remember),
           harvest = land * yield * .5 - land * sowing_rate) #%>%  # *.5 is for fallow
    #select(-yield)
}
```

```{r remember}
remember <- function(yield_memory, yield){
 # rnorm(1, yield, yield * 0.0333) %>%  #memory is fuzzy
    append(yield_memory[-length(yield_memory)], yield, after = 0) # remove the last entry in the vector and add new yield to the begining
}
```

Agents use the peak-end rule when accessing memory.
```{r}
peak_end <- function(x){
  map_dbl(x, ~mean(c(.x[1], min(.x))))
}
```

```{r}
random.series <- rnorm(100, mean = 200, sd = 50)
mem <- c()
for(i in 1:100){
  mem <- c(mem, mean(c(random.series[i], min(random.series[i:(min(i+14, 100))]))))  
}
qplot(x = 1:100, y = rev(random.series), geom = 'line') +
  geom_line(aes(y = rev(mem)), color = 'red', linetype = 2) +
  labs(x = 'Year', y = 'Crop Yield', title = 'Impact of peak-end rule on yield memory', subtitle = 'Randomly generated crop yields (black) and resulting "memory" (red), given a 15 year memory length') +
  theme_minimal()
```

Finally eat the harvested food, updating storage and food_ratio accordingly
```{r}
eat <- function(households){
  households %>%
    mutate(total.cal.req = n_inhabitants * wheat.req,
           food_ratio = pmin(1, (storage + harvest) / total.cal.req),
           old.storage = storage,
           storage = if_else(total.cal.req <= storage, harvest, pmax(harvest - (total.cal.req - old.storage), 0))) %>%
    select(-old.storage, -total.cal.req, -harvest)
}
```

### Demography

Generate fertility and mortality tables from pre-prepared data.
```{r birth-death}
fertility_table <- tibble(
  age = 10:49,
  fertility_rate = rep(c(0.022, 0.232, 0.343, 0.367, 0.293, 0.218, 0.216, 0.134), each = 5)
) %>%
  .[-1:-2,]

mortality_table <- tibble(
  age = 0:95,
  mortality_rate = c(0.4669, rep(0.0702, 4), rep(c(0.0132, 0.0099, 0.0154, 0.0172, 0.0195, 0.0223, 0.0259, 0.0306, 0.0373, 0.0473, 0.0573, 0.0784, 0.1042, 0.1434, 0.2039), each = 5), rep( 0.2654, 16))
)

cd <- demogR::cdmltw()$nmx[3,]
mortality_table_cd <- tibble(
  age = 0:99,
  mortality_rate = c(cd[1], rep(cd[2], 4), rep(cd[3:21], each = 5))
)
```

```{r echo = F}
ggplot(mortality_table, aes(age)) +
  geom_line(aes(y = mortality_rate), color = 'red') +
  geom_line(data = mortality_table_cd, aes(y = mortality_rate), linetype = 2, color = 'red') +
  geom_line(data = fertility_table, aes(y = fertility_rate), color = 'blue') +
  labs(title = 'Fertility and Mortality Rates', subtitle = 'Per capita fertility (blue) and mortality (red) in the Roman Empire', x = 'Age', y = 'Vital rate') +
  theme_bw()
```

Calculate functions for vital rate elasticities from pre-prepared data.
```{r vital-rate-elasticity, warning = F, message = F}
fertility_elasticity <- read_csv('data/fertility_data.csv', skip = 1) %>% 
  rename(food_ratio = X, fertility_reduction = Y) %>%
  gam(fertility_reduction ~ s(food_ratio, k = 35), family = betar(eps = 0.0001), dat = .)

survival_elasticity <- read_csv('data/mortality_dataset.csv', skip = 1)
survival_elasticity <- bind_rows(
    survival_elasticity[1:2] %>% mutate(age = 1) %>% rename(food_ratio = X, survival_reduction = Y), 
    survival_elasticity[3:4] %>% mutate(age = 25) %>% rename(food_ratio = X_1, survival_reduction = Y_1),
    survival_elasticity[5:6] %>% mutate(age = 5) %>% rename(food_ratio = X_2, survival_reduction = Y_2),
    survival_elasticity[7:8] %>% mutate(age = 65) %>% rename(food_ratio = X_3, survival_reduction = Y_3)) %>%
  filter(!is.na(food_ratio)) %>%
  gam(survival_reduction ~ te(food_ratio, age, k = c(45, 4)), family = betar(eps = 0.0001), dat = .)
```

For simplicity now, make survival elasticity like fertility and have a constant function for all age classes.
```{r}
survival_elasticity <- read_csv('data/mortality_dataset.csv', skip = 1) %>%
  .[3:4] %>% 
  mutate(age = 25) %>% 
  rename(food_ratio = X_1, survival_reduction = Y_1) %>%
  filter(!is.na(food_ratio)) %>%
  gam(survival_reduction ~ s(food_ratio, k = 15), family = betar(eps = 0.0001), dat = .)
```

Birth
```{r reproduction}
calculate_births <- function(households){
  households %>%
    unnest(inhabitants) %>%
    left_join(fertility_table, by = 'age') %>%  # find the fertility rate corresponding to age
    mutate(fertility_rate = if_else(is.na(fertility_rate), 0, fertility_rate),
           fertility_reduction = predict(fertility_elasticity, ., type = 'response'),
           baby = rbernoulli(n(), fertility_rate / 2 * fertility_reduction)) %>%  # divide by two to make everyone female ...
    group_by(settlement, household) %>%
    summarise(births = sum(baby)) %>%
    .$births
}

give_birth <- function(inhabitants, births){
  if(births > 0) inhabitants <- add_row(inhabitants, age = rep(0, births))
  return(inhabitants)
}

reproduce <- function(households){
  households %>%
  mutate(births = calculate_births(.),
         inhabitants = map2(inhabitants, births, give_birth)) %>%
    select(-births)
}
```

Death
```{r}
die <- function(households){
  households %>%
    unnest(inhabitants) %>%
    inner_join(mortality_table_cd, by = 'age') %>% # inner join has the effect of killing off all those over 95
    mutate(#survival_reduction = predict(survival_elasticity, ., type = 'response'),
           survive = rbernoulli(n(), (1 - mortality_rate) * 1)) %>% # survival_reduction)) %>%
    filter(survive == T) %>%
    mutate(age = age + 1) %>%
    select(-survive, -mortality_rate) %>% #, -survival_reduction) %>%
    nest(age, .key = inhabitants)
}
```

Agents first reproduce, then die, then age, and finally remove households with no inhabiants.
```{r birthdeath}
birth_death <- function(households){
 # mem <- select(households, settlement, household, yield_memory) #separate out the yield memory column and add it back in the end due to this issue: https://github.com/tidyverse/tidyr/issues/249

  households %>%
    reproduce %>%
    die %>% 
    mutate(n_inhabitants = map_int(inhabitants, nrow),
           laborers = map_dbl(inhabitants, ~ filter(.x, age >= 15 & age < 65) %>% nrow)) #%>%
    #left_join(mem, by = c('settlement', 'household'))
}
```

If households get larger than 10 inhabitants, they split into two.
```{r fission, eval=FALSE, include=FALSE}
fission <- function(households){
  static.households <- households %>% 
    filter(n_inhabitants < 10)
  
  if(nrow(static.households) == nrow(households)){
    return(households)
  } else{
  tmp <- households %>%
    filter(n_inhabitants >= 10) %>% 
    mutate(indices = map(inhabitants, ~sample.int(nrow(.x), nrow(.x) * 0.5)))
  
  old <- tmp %>% mutate(storage = .5 * storage,
                       inhabitants = map2(inhabitants, indices, ~slice(.x, .y)),
                       n_inhabitants = map_int(inhabitants, nrow))
  
  new <- tmp %>% mutate(storage = .5 * storage,
                inhabitants = map2(inhabitants, test, ~slice(.x, -.y)),
                n_inhabitants = map_int(inhabitants, nrow))
  
  return(bind_rows(static.households, old, new) %>%
    mutate(n_inhabitants = map_int(inhabitants, nrow),
           laborers = map_dbl(inhabitants, ~filter(.x, age >= 20 & age < 45) %>% nrow)))
  }
}
```


# Simulation
Start with the populaiton level parameters, i.e. how many settlements, households, and individuals to start the simulation with.

```{r pop_params}
create.households <- function(x){
  tibble(household = 1:x,
         n_inhabitants = init_inhabitants,
         storage = n_inhabitants * wheat.req,
         yield_memory = max.yield,#list(rep(max.yield, memory_length)),
         land = calc_land_req(n_inhabitants, peak_end(yield_memory)),
         food_ratio = 1) %>%
    mutate(inhabitants = map(n_inhabitants, create.inhabitants),
           laborers = map_dbl(inhabitants, ~filter(.x, age >= 15 & age < 65) %>% nrow))
}

create.inhabitants <- function(x){
  tibble(age = round(runif(x, 1, 60)))
}
```

Create a 2 villages of 10 households with 6 people in each.
```{r}
init_settlements <- 1 
init_households <- 1
init_inhabitants <- 5

population <- tibble(settlement = 1:init_settlements,
                     x = 1, y = 1,
                     households = init_households) %>%
              mutate(households = map(households, create.households))
```

Make an environment data table. For now we'll keep it to dummy value for development and testing purposes, see the following section for a "real environment".

```{r}
environment <- tibble(x = 1, y = 1, area = 1, arable_proportion = 1, precipitation = 1, vegetated_fraction = 1, vegetation_stage = 50)
```

```{r}
# this function takes a settlement (tibble of households), and makes the households do stuff
household_dynamics <- function(settlements){
  unnest(settlements) %>%
    allocate_land %>%
    farm %>%
    eat %>%
    birth_death %>%
    nest(household:last(everything()), .key = households)
}

# this function takes the master dataframe and calculates the per patch climatic yields from per patch precipitation
environmental_dynamics <- function(settlements){
  settlements %>%
    mutate(climatic_yield = calc_climatic_yield(precipitation))
}
```

Calculate population totals for plotting
```{r}
calc_pop <- function(settlements){
  settlements %>% 
    unnest(households) %>%
    .$n_inhabitants %>%
    sum
}
calc_food <- function(settlements){
  settlements %>%
    unnest(households) %>%
    .$food_ratio %>%
    mean
}
```


```{r simulation}
nsim <- 400
replicates <- 4

sim.out <- pblapply(1:replicates, cl = 4, function(j){
  test <- population %>% left_join(environment, by = c('x', 'y'))
  sim.out.tmp <- tibble(year = rep(0, nsim + 1), population = c(calc_pop(population), rep(0, nsim)), food_ratio = 1, replicate = j)
  for(i in 1:nsim){
        sim.out.tmp[i + 1,] <- c(i, 0, 0, j)
    if(nrow(test %>% unnest) > 0){
          sim.out.tmp[i + 1,] <- c(i, calc_pop(test), calc_food(test), j)
  test <- test %>% 
    environmental_dynamics %>%
    household_dynamics
  }
  }
  return(sim.out.tmp)
}) %>% bind_rows
```

```{r sim-plots}
ggplot(sim.out, aes(year, population, group = replicate)) +
  geom_smooth(aes(color = as.factor(replicate))) +
  geom_line(alpha = .1) +
  theme_minimal()

ggplot(sim.out, aes(year, food_ratio, group = replicate)) +
  geom_smooth(aes(color = as.factor(replicate))) +
  geom_line(alpha = .1) +
  theme_minimal()
```

```{r}
sim.out %>%
  filter(year == nsim) %>%
  ggplot(aes(x = population)) +
  geom_density() +
  theme_minimal()
sim.out %>%
  filter(year == nsim) %>%
  ggplot(aes(x = food_ratio)) +
  geom_density() +
  theme_minimal()
```

# Environment

## Input data

Set a bounding box for the study area.
```{r bbox}
bbox <- extent(5, 11.5, 34, 37.5)
elevation <- raster('~/gdrive/Data/SRTM_1km.tif') %>% crop(bbox)
```

```{r climate}
prec <- brick('data/CHELSA/prec_1km.tif') %>% 
  sum %>%
  projectRaster(elevation)
```

Calculate slope in degrees from 90m SRTM elevation (90m data from http://www.earthenv.org/DEM).
```{r slope}
elev.list <- list.files('data/topography/', pattern = 'bil$', full.names = T, recursive = T) %>%
  map(raster)
elev.list$fun <- mean

slope <- do.call(mosaic, elev.list) %>% 
  crop(bbox) %>%
  terrain(opt = 'slope', unit = 'degrees')
```

Where is the arable land? Calculate the proportion of land with =< 5 degree slope. Also find land with slope between 5 and 15 degrees, as potential land for terracing.
```{r arable_land}
arable_proportion <- (slope <= 5) %>% 
  aggregate(fact = 10, fun = sum) %>% 
  mask(elevation) %>%
  `/`(100)
```


Get the area of each grid cell, in km2.
```{r}
area <- area(elevation)
```

Let's put all these environmental rasters together in a single brick for easier access.
```{r}
environment <- brick(arable_proportion, prec, area) %>%
  as.data.frame(na.rm = T, xy = T, centroids = T) %>%
  as_tibble %>%
  remove_rownames %>%
  rename(arable_proportion = layer.1, precipitation = layer.2, area = layer.3) %>%
  mutate(precipitation = precipitation / 1000,
         x = round(x, 6), y = round(y, 6))
environment
```
 
 Now import archaeological data
```{r}
settlement_locations <- pl_search_places() %>%
  filter(locationPrecision == 'precise') %>%
  select(title, type = featureTypes, description, 
         lon = reprLong, lat = reprLat, 
         periods = timePeriods, minDate, maxDate, id, tags) %>%
  filter(between(lat, 34, 37.5) & between(lon, 5, 11.5)) %>%
  collect  %>%
  filter(str_detect(periods, 'R')) %>%
  filter(str_detect(type, 'settlement'))
```
 
```{r echo = F}
ggplot(settlement_locations, aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1) +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void()
```

```{r}
init_households <- 1
init_inhabitants <- 4

population <- settlement_locations %>%
  select(lon:lat) %>%                        # get the x and y coordinates of settlements
  rasterize(prec, fun = 'count') %>%    # create a raster representing the count of settlements in each cell
  mask(prec) %>%
  as.data.frame(xy = T, na.rm = T, centroids = T) %>%
  remove_rownames %>%
  as.tibble %>%
  mutate(settlement = 1:n(), x = round(x, 6), y = round(y, 6)) %>%   # the coordinates here and in the environment raster need to be rounded to correct for floating point inequality issues
  select(-layer) %>%
  mutate(households = init_households,
         households = map(households, create.households))
```

```{r}
nsim <- 400

test <- left_join(population, environment) %>%
  filter(!is.na(precipitation)) %>%
  filter(arable_proportion > 0)
test %>% 
  mutate(groups = rep(1:3, length.out = nrow(.))) %>%
  split(.$groups)


system.time(
for(i in 1:nsim){
  test <- test %>% 
    environmental_dynamics %>%
    household_dynamics
  }
)
```

```{r}
ggplot(test %>% unnest, aes(x,y)) +
geom_point(aes(size = n_inhabitants)) +
    coord_quickmap()
```

```{r}
```

Functions for converting vegetation succession stage to woody biomass (per m2) and back again.
```{r}
vegetation_to_biomass <- function(vegetation){
  ifelse(vegetation > 35, vegetation * 0.08066667 - 2.083333, 
         ifelse(vegetation > 18, vegetation * 0.004705882 + 0.5752941, 
                ifelse(vegetation > 7, vegetation * 0.05090909 - 0.2563636, 
                       vegetation * 0.01428571)))
}
biomass_to_vegetation <- function(biomass){
  ifelse(biomass > 0.74, biomass * 12.39669 + 25.82645, 
         ifelse(biomass > 0.66, biomass * 212.5 - 122.25, 
                ifelse(biomass > 0.1, biomass * 19.64286 + 5.035714, 
                       biomass * 70)))
}

qplot(0:50, vegetation_to_biomass(0:50), geom = 'line') +
  labs(title = 'Vegetation to biomass conversion', x = 'Vegetation succession stage', y = 'Biomass (kg/m2)') +
  theme_minimal()
```



```{r}
wood_req <- 1600 # yearly firewood requirement (kg) per person
max_wood_gather_intensity <- 0.08 * 1e6 # maximum wood gathering intensity in kg/m2, converted to to kg/km2

gather_wood <- function(settlements){
  settlements %>%
    mutate(pop = map_dbl(households, ~sum(.$n_inhabitants)),
           available_biomass = vegetation_to_biomass(vegetation_stage) * vegetated_fraction * 1e6,  # the available biomass in the entire vegetated fraction of the cell
           new_biomass = available_biomass - wood_req * pop ^ 0.8,
           vegetation_stage = biomass_to_vegetation(new_biomass / (vegetated_fraction * 1e6))
           ) %>%
    select(-pop, -available_biomass, -new_biomass) %>%
    mutate(vegetation_stage = pmin(vegetation_stage + 1, max_veg)) # regenerate one succession unit of vegetation a year, up to the maximum vegetation
}
```

```{r}
#this plot says that a population of less than 130 people can get their wood requirmenets from a fully vegetated 1km2 grid cell at the gathering intensity rate of 0.08 kg/m2. Above that population, right now people just extract more wood than they should ethnogrpahically. ideally at this threshold they would start gathering wood from naighboring patches.
qplot(1:200, (wood_req * (1:200) ^ 0.8), geom= 'line') +
  geom_hline(yintercept = max_wood_gather_intensity)
```

### LCPS
```{r eval=FALSE, include=FALSE}
altDiff <- function(x){x[2] - x[1]}
hd <- transition(elevation, altDiff, 16, symm=FALSE)
adj <- adjacent(elevation, cells=1:ncell(elevation), pairs=TRUE, directions=16)
slope.c <- geoCorrection(hd, type = 'c')
speed.c <- slope.c
speed.c[adj] <- 6 * exp(-3.5 * abs(slope.c[adj] + 0.05))
Conductance.c <- geoCorrection(speed.c, type = 'c')
rm(slope.c, speed.c, hd, adj)
```

```{r}
test <- settlement_locations %>%
  select(lon:lat) %>%
  mutate(costs = map2(lon, lat, get_farming_catchment))

get_farming_catchment <- function(x, y, conductance = Conductance.c){
  accCost(conductance, c(x, y)) %>%
    as.data.frame(xy = T, na.rm = T) %>%
    rename(distance = layer) %>%
    mutate(distance = distance / 3600) %>%
    filter(distance <= .5)
}
```

