---
title: "analysis"
author: "Nick Gauthier"
date: "March 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(sf)
library(tidygraph)
library(ggraph)
library(patchwork)
```


# real world
```{r}
#bbox_ll <- extent(5, 11.5, 34, 37.5)
#bbox_utm <- extent(190000, 735000, 3780000, 4150000) # this goes into algeria
bbox_ll <- extent(7, 11.5, 34, 37.5)
bbox_utm <- extent(345000, 735000, 3780000, 4150000)
```

```{r}
prec <- brick('data/CHELSA/prec_1km.tif') %>%
  .[[c(1:5, 10:12)]] %>% # extract october -- may
  sum %>% 
  `/`(1000) %>%
  projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m") %>%
  crop(bbox_utm)
```

```{r, cache = TRUE}
boundary <- (!prec) %>% 
  rasterToPolygons(dissolve = TRUE) %>% 
  st_as_sf %>%
  st_cast('POLYGON') %>%
  mutate(area = st_area(geometry)) %>%
  filter(area == max(area))
```

```{r, cache = TRUE}
slope <- raster('data/topography/earthenv_90m.tif') %>% 
  crop(bbox_ll) %>%
  projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m") %>%
  terrain(opt = 'slope', unit = 'degrees') %>%
  crop(bbox_utm) %>%
  mask(as(boundary, 'Spatial')) 
```

```{r}
max_veg_stage <- raster('data/MODIS/veg_stage.tif') %>%
    projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m", method = 'ngb')
```

```{r}
prec_200 <- brick('data/CCSM4/prec_annual_1km.tif') %>%
   crop(bbox_ll) %>%
  `/`(1000) %>%
  projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m", method = 'ngb') %>%
  crop(bbox_utm) %>%
  mask(as(boundary, 'Spatial'))
```

data from 
Barbarossa, V et al. figshare https://dx.doi.org/10.6084/m9.figshare.c.3890224
```{r}
streamflow <- brick('data/topography/FLO1K.5min.ts.1960.2015.nc/FLO1K.5min.ts.1960.2015.qav.nc') %>%
  crop(bbox_ll) %>%
  mean(na.rm = T) %>%
  projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m") %>%
  crop(bbox_utm) %>%
  mask(as(boundary, 'Spatial')) 

plot(streamflow)
```
```{r}
runoff <- list.files('~/Downloads/TerraClimate/', full.names = TRUE, pattern = '*.nc') %>%
  map(brick) %>%
  map(~crop(., bbox_ll)) %>%
  map(~sum(.)) %>%
  brick %>%
  mean %>%
  projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m") %>%
  crop(bbox_utm) %>%
  mask(as(boundary, 'Spatial')) 

plot(runoff)
```

```{r}
m1 <- prec %>%
  projectRaster(runoff) %>%
  brick(runoff, .) %>%
  as.data.frame(xy = TRUE, na.rm = TRUE) %>%
  rename(runoff = layer.1, prec = layer.2) %>%
  as.tbl() %>%
  mgcv::gam(runoff ~ s(prec), data = ., family = Gamma(link = 'log'), method = 'REML', select = TRUE)

runoff_200 <- prec_200 %>%
  mean %>%
  setNames('prec') %>%
  predict(m1, type = 'response') %>%
  projectRaster(runoff)
```

```{r}
hex <- boundary %>%
  st_make_grid(cellsize = 7000, square = FALSE) %>%
  st_sf %>%
  mutate(precipitation = raster::extract(prec, ., fun = mean, na.rm = TRUE),
         area = as.numeric(st_area(geometry)) * 1e-6, # convert from m2 to km2
         arable = raster::extract(slope <= 5, ., fun = sum, na.rm = TRUE) / raster::extract(setValues(slope, 1), ., fun = sum, na.rm = TRUE),
         cultivable_area = area * arable,
         max_veg_stage = raster::extract(max_veg_stage, ., fun = modal, na.rm = TRUE),
         streamflow = raster::extract(streamflow, ., fun = mean, na.rm = TRUE),
         runoff = raster::extract(runoff, ., fun = mean, na.rm = TRUE),
         precipitation_200 = raster::extract(mean(prec_200), ., fun = mean, na.rm = TRUE),
        runoff_200 = raster::extract(runoff_200, ., fun = mean, na.rm = TRUE)) %>%
  filter(cultivable_area >= 1) # need at least 1km2
```


```{r}
e1 <- ggplot(hex) +
geom_sf(aes(fill = precipitation * 1000, color = precipitation * 1000)) +
  scale_fill_distiller(palette = 'YlGnBu', direction = 1, name = 'Total (mm)') +
    scale_color_distiller(palette = 'YlGnBu', direction = 1, name  = 'Total (mm)') +
  labs(subtitle = 'Annual precipitation') +
  coord_sf(datum = NA) +
  theme_void()

e2 <- ggplot(hex) +
  geom_sf(aes(fill = cultivable_area, color = cultivable_area)) +
    scale_fill_viridis_c(name = 'Area (km2)') +
  scale_color_viridis_c(name = 'Area (km2)') +
  labs(subtitle = 'Arable land') +
  coord_sf(datum = NA) +
  theme_void()

lc_colors <- c('#fffbc3', '#dcce00', '#97bf47', '#018200')
lc_labels <- c('barren', 'grass', 'shrub', 'woodland')
e3 <- hex %>%
  mutate(Stage = as.factor(max_veg_stage)) %>%
ggplot() +
  geom_sf(aes(fill = Stage, color = Stage)) +
  scale_fill_manual(values = lc_colors, labels = lc_labels) +
  scale_color_manual(values = lc_colors, labels = lc_labels) +
  labs(subtitle = 'Potential natural vegetation') +
  coord_sf(datum = NA) +
  theme_void()

e4 <- ggplot(hex) +
  geom_sf(aes(fill = streamflow, color = streamflow)) +
  coord_sf(datum = NA) +
  scale_fill_distiller(palette = 'BrBG', direction = 1, name = 'Volume (m3/s)') +
    scale_color_distiller(palette = 'BrBG', direction = 1, name = 'Volume (m3/s)') +
  labs(subtitle = 'Streamflow') +
  theme_void()

e1 + e2 + e3 + e4 + plot_annotation(title = 'Environmental inputs')
ggsave('environmental_inputs.pdf', height = 8, width = 12)
```

```{r}
write_sf(hex, 'NA_env.shp')

hex %>%
  mutate(streamflow_annual = streamflow * 3.154e+7,
        runoff = streamflow_annual / (cultivable_area * 1000000)) %>%
ggplot() +
  geom_sf(aes(fill = runoff > .5, color = runoff>.5)) +
    scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_sf(datum = NA) +
  theme_void()
```


```{r}
tbl_graph(nodes = read_csv('data/orbis_nodes_0514.csv'),
          edges = read_csv('data/orbis_edges_0514.csv'))
```


```{r}
hyde <- raster('../North-Africa/cropland200AD.asc') %>%
  `crs<-`(value = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0') %>%
    projectRaster(crs = "+proj=utm +zone=32 +north +ellps=WGS84 +datum=WGS84 +units=m") %>%
  crop(bbox_utm)

plot(hyde)
```


## simulation results

```{r}
sim1 <- readRDS('experiments/eqpop_nomigration_noirrigation')
sim2 <- readRDS('experiments/eqpop_nomigration_irrigation')
sim3 <- readRDS('experiments/eqpop_nomigration_noirrigation_CE200')
sim4 <- readRDS('experiments/eqpop_nomigration_irrigation_CE200')
```

```{r fig.width = 9, fig.height = 9}
sim_dat <- rbind(sim1, sim2, sim3, sim4) %>%
  mutate(irrigation = as.factor(rep(rep(c('No Irrigation', 'Runoff Harvesting'), each = 2453), 2)),
          time = as.factor(rep(c('Present day', '200 CE'), each = 4906)),
         population = map_dbl(sim, ~.$population)) %>%
    select(-sim) %>% 
  mutate(population = na_if(population, 0))

population_annotation <- sim_dat %>%
  group_by(irrigation, time) %>%
  summarise(population = round(sum(population, na.rm = TRUE) / 1000000, 1)) %>%
  mutate(population = paste('Total:', population, 'million'))

ggplot(sim_dat) +
  geom_sf(aes(fill = population), color = 'grey60', lwd = .05) +
  coord_sf(datum = NA) +
  facet_grid(irrigation ~ time) +
  scale_fill_viridis_c(na.value = 'white') +
  ggtitle('Equilbirium population after 600 years', 'With and without runoff irrigation, 200 CE and present day') +
  geom_text(data = population_annotation, aes(x = 500000, y = 3800000, label = population)) +
  theme_void()

ggsave('test.pdf', height = 9, width = 9)
```


# Environment

## Input data

Set a bounding box for the study area.
```{r bbox}
bbox <- extent(5, 11.5, 34, 37.5)
elevation <- raster('~/gdrive/Data/SRTM_1km.tif') %>% crop(bbox)
```


```{r eval=FALSE, include=FALSE}
prec_200 <- (brick('data/CCSM4/trans1-850AD.cam.h0.PRECC.010001-019912.nc') + 
             brick('data/CCSM4/trans1-850AD.cam.h0.PRECL.010001-019912.nc')) %>%
  rotate %>%
  crop(bbox, snap = 'out') %>% 
  `*`(2.628e+9) %>% 
  projectRaster(prec) %>%
  crop(bbox) %>%
  mask(prec) %>% 
  stackApply(rep(1:100, each = 12), sum)
gc()
prec_2000 <- (brick('data/CCSM4/trans1-850AD.cam.h0.PRECC.190001-199912.nc') + 
              brick('data/CCSM4/trans1-850AD.cam.h0.PRECL.190001-199912.nc')) %>%
  rotate %>%
  crop(bbox, snap = 'out') %>% 
  `*`(2.628e+9) %>% 
  projectRaster(prec) %>%
  crop(bbox) %>%
  mask(prec) %>% 
  stackApply(rep(1:100, each = 12), sum) %>%
  mean
gc()
prec_annual_1km <- (prec * prec_200 / prec_2000) / 1000
#writeRaster(prec_annual_2k, 'data/CCSM4/prec_annual_1km.tif')
rm(prec, prec_200, prec_2000)
```

Calculate slope in degrees from 90m SRTM elevation (90m data from http://www.earthenv.org/DEM).
```{r slope}
elev.list <- list.files('data/topography/', pattern = 'bil$', full.names = T, recursive = T) %>%
  map(raster)
elev.list$fun <- mean

slope <- do.call(mosaic, elev.list) %>% 
  crop(bbox) %>%
  terrain(opt = 'slope', unit = 'degrees')
```

Where is the arable land? Calculate the proportion of land with =< 5 degree slope. Also find land with slope between 5 and 15 degrees, as potential land for terracing.
```{r arable_land}
arable_proportion <- (slope <= 5) %>% 
  aggregate(fact = 10, fun = sum) %>% 
  mask(elevation) %>%
  `/`(100)
```




Let's put all these environmental rasters together in a single brick for easier access.
```{r}
environment <- brick(arable_proportion, runoff, area, max_veg_stage)
names(environment) <- c('arable_proportion', 'runoff', 'area', 'max_veg_stage')
plot(environment)
```
 
 Now import archaeological data
```{r}
library(pleides)
settlement_locations <- pl_search_places() %>%
  filter(locationPrecision == 'precise') %>%
  select(title, type = featureTypes, description, 
         lon = reprLong, lat = reprLat, 
         periods = timePeriods, minDate, maxDate, id, tags) %>%
  filter(between(lat, 34, 37.5) & between(lon, 5, 11.5)) %>%
  collect  %>%
  filter(str_detect(periods, 'R')) %>%
  filter(str_detect(type, 'settlement'))
```
 
```{r echo = F}
ggplot(settlement_locations, aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1) +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void()
```
