---
title: "Input data comparison"
output:
  html_document: default
  html_notebook: default
---

We have multiple options for the climate and soil data to use for the potential cultivable land analysis. Let's compare them.

```{r}
library(raster)
library(rasterVis)
library(tidyverse)
library(pbapply)
```

```{r}
bbox <- extent(-10, 20, 28, 38)
```


# Climate
First import the modis/chirps data at 3' resolution
```{r}
get_clim <- function(x){
  list.files(x, full.names = T) %>%
    stack %>%
    crop(bbox)
}

prec.mod <- get_clim('~/gdrive/Data/MOD11C3v6.0-CHIRPSv2.0_MONTHLY_03m/prec/')
temp.mod <- get_clim('~/gdrive/Data/MOD11C3v6.0-CHIRPSv2.0_MONTHLY_03m/temp/')

prec.mod[prec.mod[] == -9999] <- NA
temp.mod[temp.mod[] == -9999] <- NA
```

Next import the CHELSA data and resample to 3' resolutioin.
```{r}
# the list.files import metho for chelsa mixes up the month orders ... fix manually
prec.chelsa <- get_clim('~/Downloads/CHELSA/prec/') %>%
  aggregate(fact = 6) %>%
  .[[c(1,5:12, 2:4)]]
temp.chelsa <- get_clim('~/Downloads/CHELSA/temp/') %>% 
  `/`(10) %>%
  aggregate(fact = 6) %>%
  .[[c(1,5:12, 2:4)]]
```


Compare the temp maps
```{r}
levelplot(brick(c(temp.mod, temp.chelsa))[[c(1,13,4,16,7,19,10,22)]])
levelplot(temp.mod - temp.chelsa, par.settings = BuRdTheme(), at = seq(-15,15,1))
levelplot(temp.chelsa)
```
consistent warm bias in modis lst during hotter months (makes sense if we're talking skin temp vs air temp).
consistent cold bias along coasts in winter

Compare the prec maps.
```{r}
levelplot(brick(c(prec.mod, prec.chelsa))[[c(1,13,4,16,7,19,10,22)]])
levelplot(prec.mod - prec.chelsa, par.settings = PuOrTheme(), at = seq(-260, 260, 10))
```

In the end, it looks like the modis/chirps dataset has too many artifacts, even though its nicer.

## Growing degree days
```{r cache = T}
gdd <- function(temp.map){
  
  temp.daily <- temp.map[[rep(1, 365)]]
  
  temp.map %>%
    getValues %>%
    pbapply(1, function(x){
    if(any(is.na(x))){return(rep(NA, 365))} else{
      dat <- data.frame(doy =  c(15, 46, 74, 105, 135, 166, 196, 227, 258, 288, 319, 349),
                        temp = x)
      mod <- gam(temp ~ s(doy, bs = 'cc', k = 12), knots = list(doy = c(1, 366)), data = dat)
      return(predict.gam(mod, data.frame(doy = 1:365)))
    }
}) %>% 
    t %>%
  setValues(temp.daily, .) %>% 
    reclassify(c(-Inf, 5, 0)) %>%
    sum
}

gdd.mod <- gdd(temp.mod)
gdd.chelsa <- gdd(temp.chelsa)
levelplot(brick(c(gdd.mod, gdd.chelsa)))
gdd.index <- 1 / (1 + exp(0.0052 * (1334 - gdd.chelsa)))
```

```{r}
library(SPEI)
lats <- temp.mod[[1]] %>% coordinates %>% .[,2] %>% setValues(temp.mod[[1]], .)

pet.mod <- thornthwaite((temp.mod %>% getValues %>% t), lat = getValues(lats), na.rm = T) %>%
  t %>%
  setValues(temp.mod, .)
pet.chelsa <- thornthwaite((temp.chelsa %>% getValues %>% t), lat = getValues(lats), na.rm = T) %>%
  t %>%
  setValues(temp.chelsa, .)


levelplot(brick(c(pet.mod, pet.chelsa))[[c(1,13,4,16,7,19,10,22)]])
levelplot(pet.mod - pet.chelsa, par.settings = PuOrTheme())
```
calculate alpha
```{r}
alpha <- raster('~/Downloads/CGIAR PET/ALPHA/alpha/') %>% 
  crop(bbox) %>%
  aggregate(fact = 6) %>%
  `/`(100)
levelplot(alpha)
alpha.index <- 1 / (1 + exp(14.705 * (0.3295 - alpha)))
levelplot(alpha.index)
```

# Soil Data
the polygon for the hwsd data make too large artifacts for this scale
```{r}
ph.hwsd <- raster('~/Downloads/Harmonised Soil Database/HWSD_1247/data/T_PH_H2O.nc4') %>%
  crop(bbox)
c.hwsd <- raster('~/Downloads/Harmonised Soil Database/HWSD_1247/data/AWT_T_SOC.nc4') %>%
  crop(bbox)
ph.sg <- raster('~/Downloads/SoilGrids250/PHIHOX_M_sl4_250m_ll.tif') %>%
  crop(bbox) %>%
  aggregate(fact = 24) %>%
  `*`(.1)
c.sg <- raster('~/Downloads/SoilGrids250/OCSTHA_M_30cm_250m_ll.tif') %>%
  crop(bbox) %>%
  `*`(.1) %>%
  aggregate(fact = 24)
levelplot(brick(c(c.sg, c.hwsd)))
levelplot(brick(c(ph.sg, ph.hwsd)))

```

check indices from ramankutty
```{r}
a <- 3.9157; b <- 1.3766; c <- 3.468; d <- -.0791; e <- -27.33
c.index <- (a / (1 + exp(b * (c - c.sg)))) * (a / (1 + exp(d * (e - c.sg))))
levelplot(c.index)

ph.index <- ph.sg
ph.index[ph.sg < 6.5] <- -2.085 + .475 * ph.sg[ph.sg < 6.5]
ph.index[ph.sg >= 6.5 & ph.sg < 8] <- 1
ph.index[ph.sg >=8] <- 1 - 2 * ph.sg[ph.sg >=8]

levelplot(ph.sg)
levelplot(ph.index)
```

```{r}
soil.index <- (ph.index * c.index) %>% reclassify(c(-Inf, 0, 0))
levelplot(soil.index)
```
```{r}
climate.index <- (gdd.index * alpha.index)
levelplot(climate.index)
```

```{r}
levelplot(climate.index * soil.index)
```
```{r}
gplot(climate.index * soil.index) +
  geom_raster(aes(fill = value)) +
  geom_point(data = villas, aes(lon, lat)) +
  coord_quickmap()
```

