---
title: "fiscal-demographic"
author: "Nick Gauthier"
date: "January 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=F, warning=F}
library(deSolve)
library(phaseR)
```

Second let's write a convenience function that calls on phaseR under the hood to generate the flow field, nullclines, and sample trajectories for a given system and parameterization.

```{r}
phasePlot <- function(mod, params, xmax = 1, ymax = 1){
  xlim <- c(.1, xmax)
  ylim <- c(0, ymax)
  
  y0 <- matrix(c(1, 2),
             ncol = 2, nrow = 1,
             byrow = TRUE)  
  
  flw <- flowField(mod, xlim = xlim, ylim = ylim, parameters = params, 
                   xlab = 'Population', ylab = 'State coffers', add = F)
  nc <- nullclines(mod, xlim = xlim, ylim = ylim, parameters = params)
  trj <- trajectory(mod, y0 = y0, tlim = c(0,200), col = rep('black', nrow(y0)), parameters = params)
}
```

# Model 1: Simple consumer-resource system with network effects
First, we'll replicate the model of Muneepeerakul and Qubbaj (2012). It's a simple consumer resource system, with parameterized flows of population and resources (i.e. immigration and trade).

Setup the model.

```{r}
netMod <- function(t, y, parameters){
    r <- parameters[1]
    q <- parameters[2]
    N <- y[1]
    K <- y[2]

    dy <- numeric(2)
    dy[1] <- r * N * (1 - N / K)
    dy[2] <- (q / (q - 1)) * N / (N + (1/(q - 1))) - K
    list(dy)
}

derivative <- function(t, y, parameters) {
  alpha <- parameters[1]
  beta  <- parameters[2]
  x     <- y[1]
  y     <- y[2]
  dy    <- numeric(2)
  dy[1] <- alpha*y
  dy[2] <- beta*x
  list(dy)
}
```

No scaling
```{r fig.width = 6, fig.height = 6}
phasePlot(netMod, c(0.02, 1.2), xmax = 1.5, ymax = 1.5)

flowField(netMod, xlim = c(.1, 5), ylim = c(0, 5), parameters = c(0.02, 1.2), 
                   xlab = 'Population', ylab = 'State coffers', add = F)

phasePlaneAnalysis(netMod, c(.01,1.5), c(0,1.5), parameters = c(0.2, 1.2))
```