author: "Nick Gauthier"
date: 
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and setup
Load required packages
```{r}
library(raster)
library(rgdal)
library(rasterVis)
library(viridis)
library(tidyverse)
library(stringr)
library(pleiades)
```
Define map extent.
```{r}
bbox <- extent(5, 11.5, 34, 37.5)
```
## Elevation Data
Get country boundaries for spain, portugal, and italy to mask out from elevation
```{r}
euro.mask <- rbind(getData('GADM', country = 'PRT', level = 0),
                   getData('GADM', country = 'ESP', level = 0),
                   getData('GADM', country = 'ITA', level = 0))
```

First download and mosaic 90m srtm tiles over the study area.
```{r eval = F}
#getData('SRTM', lon = -8, lat = 35)
#getData('SRTM', lon = -3, lat = 35)
#getData('SRTM', lon = 2, lat = 35)
getData('SRTM', lon = 7, lat = 35)
getData('SRTM', lon = 12, lat = 35, path = 'data/')
#getData('SRTM', lon = -8, lat = 40)
#getData('SRTM', lon = -3, lat = 40)
#getData('SRTM', lon = 2, lat = 40)
getData('SRTM', lon = 7, lat = 40, path = 'data')
getData('SRTM', lon = 12, lat = 40, path = 'data')



srtm <- list.files(pattern = '*.tif', recursive = T) %>% 
  lapply(raster) %>% 
  do.call(merge, .) #%>%
  mask(euro.mask, inverse = T)

writeRaster(srtm, 'srtm_na.tif')
srtm <- raster('data/srtm_na.tif')
```


```{r}
srtm <- raster('~/gdrive/Data/SRTM_1km.tif') %>% crop(bbox)
plot(srtm)
```

```{r}
plot(srtm)
```

```{r eval = F}
slope <- terrain(srtm, opt = 'slope', filename = 'slope.tif', overwrite = T)
aspect <- terrain(srtm, opt = 'aspect', filename = 'aspect.tif', overwrite = T)
TPI <- terrain(srtm, opt = 'TPI', filename = 'TPI.tif', overwrite = T)
TRI <- terrain(srtm, opt = 'TRI', filename  = 'TRI.tif', overwrite = T)
roughness <- terrain(srtm, opt = 'roughness', filename = 'roughness.tif', overwrite = T)
flowdir <- terrain(srtm, opt = 'flowdir', filename = 'flowdir.tif', overwrite = T)
```

```{r eval = F}
plot(slope)
plot(aspect)
plot(TPI)
plot(TRI)
plot(roughness)
plot(flowdir)
```

```{r cache = T, eval = F}
dist_coast <- srtm %>%
  reclassify(c(-Inf, Inf, NA, NA, NA, 1)) %>% # reverse NA and non-NA cells
  distance(doEdge = T) %>% # calculate the distances
  mask(srtm) %>% # mask out ocean cells  
  `/`(1000) # convert to km
writeRaster(dist_coast, 'data/topography/coast.tif')
```

```{r}
dist_coast <- raster('data/topography/coast.tif')
plot(dist_coast)
```

```{r}
cities <- read.csv('data/NA_cities.csv') %>%
  group_by(BA_rank) %>% 
  mutate(size_est = ifelse(is.na(Size), mean(Size, na.rm=TRUE), Size)) %>%
  ungroup %>%
  mutate(population = 41.834 * Size^1.3361, pop_est = 41.834 * size_est^1.3361) %>%
  filter(between(lat, ymin(srtm), ymax(srtm)) & between(lon, xmin(srtm), xmax(srtm))) %>%
  mutate(coastal = (1 >= raster::extract(dist_coast, cbind(lon, lat))))

#coordinates(cities) <- ~lon + lat
#shapefile(cities, filename = 'NA_cities')


```



```{r}
roads <- readOGR('data/Shapefiles/ba_roads.shp') %>%
  spTransform(CRS('+proj=longlat')) %>%
  crop(bbox) %>%
  fortify

aqueducts <- readOGR('data/Shapefiles/aqueducts.shp') %>%
  spTransform(CRS('+proj=longlat')) %>%
  crop(bbox)

centuriation <- readOGR('data/Shapefiles/centuriation_polygons.shp') %>%
  spTransform(CRS('+proj=longlat')) %>%
  crop(bbox) %>%
  fortify
```

```{r}
ggplot(cities, aes(lon, lat)) +
  geom_path(data = roads, aes(long, lat, group = group), color = 'lightgrey') +
  geom_point(aes(size = pop_est * 10)) +
  scale_size_area() +
  theme_void() +
  coord_fixed() + theme(legend.position="none")

ggsave('roman_network.png')
```
```{r eval = F}
gplot(roughness) +
  geom_tile(aes(fill = value)) +
    geom_path(data = roads, aes(long, lat, group = group), alpha = .5, color = 'lightgrey') +
  scale_fill_viridis() +
  coord_equal()
```

```{r}
places <- pl_search_places() %>%
  filter(locationPrecision == 'precise') %>%
  select(title, type = featureTypes, description, 
         lon = reprLong, lat = reprLat, 
         periods = timePeriods, minDate, maxDate, id, tags) %>%
  collect  %>%
  filter(between(lat, ymin(srtm), ymax(srtm)) & between(lon, xmin(srtm), xmax(srtm))) %>%
  filter(str_detect(periods, 'R')) %>%
  print
```
Plot everything
```{r}
places %>% filter(str_detect(type, 'settlement|villa')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, alpha = .8) +
  geom_path(data = roads, aes(long, lat, group = group), color = 'lightgrey') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void() +
  geom_point(data = cities, aes(lon, lat, size = pop_est)) +
  scale_size_area()
```

Settlements
```{r}
places %>% filter(str_detect(type, 'settlement')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1, alpha = .8) +
  geom_path(data = roads, aes(long, lat, group = group), color = 'lightgrey') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void() +
  geom_point(data = cities, aes(lon, lat, size = pop_est)) +
  scale_size_area()
```
Villas
```{r}
places %>% filter(str_detect(type, 'villa')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1, alpha = .8) +
  geom_path(data = roads, aes(long, lat, group = group), color = 'lightgrey') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void() +
  geom_point(data = cities, aes(lon, lat, size = pop_est)) +
  scale_size_area()
```

Mines
```{r}
places %>% filter(str_detect(type, 'mine')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1) +
  geom_path(data = roads, aes(long, lat, group = group), alpha = .5, color = 'white') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void()
```


Springs and oases
```{r}
places %>% filter(str_detect(type, 'spring')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1) +
  geom_path(data = roads, aes(long, lat, group = group), alpha = .5, color = 'white') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void()

places %>% filter(str_detect(type, 'oasis')) %>%
ggplot(aes(lon, lat)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = F, n = 500, h = 1) +
  geom_path(data = roads, aes(long, lat, group = group), alpha = .5, color = 'white') +
  scale_fill_viridis() +
  coord_fixed() +
  theme_void()
```



Reproject and save out shapefile
```{r eval = F}
settlements <- places %>% filter(str_detect(type, 'settlement'))
coordinates(settlements) <- ~lon + lat
proj4string(settlements) <- '+proj=longlat'
settlements <- spTransform(settlements, CRS("+proj=lcc +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs"))

shapefile(settlements, filename = 'data/netlogo/settlements')
```


Compare to environmental covariates
```{r, eval=FALSE, include=FALSE}
srtm <- raster('data/SRTM_NE_250m_TIF/SRTM_NE_250m.tif') %>%
  crop(bbox)

slope <- terrain(srtm, opt = 'slope', unit = 'degrees')
aspect <- terrain(srtm, opt = 'aspect', unit = 'degrees')
TPI <- terrain(srtm, opt = 'TPI')
TRI <- terrain(srtm, opt = 'TRI')
roughness <- terrain(srtm, opt = 'roughness')
hill <- hillShade(slope, aspect, 40, 270)

topo <- brick(c(srtm, slope, aspect, TPI, TRI, roughness))
rm(srtm, slope, aspect, TPI, TRI, roughness)

plot(slope<9)
```


```{r cache = T}
library(gdistance)
altDiff <- function(x){x[2] - x[1]}
hd <- transition(srtm, altDiff, 16, symm=FALSE)
cell.targets <- Which(!is.na(srtm), cells = T)
adj <- adjacent(srtm, cells=cell.targets, target = cell.targets, pairs=TRUE, directions=16)
```

Create conductance matrix with geocorrection for lcps
```{r lcp_conductance, cache = T}
slope.c <- geoCorrection(hd, type = 'c')
speed.c <- slope.c
speed.c[adj] <- 6 * exp(-3.5 * abs(slope.c[adj] + 0.05))
Conductance.c <- geoCorrection(speed.c, type = 'c')
rm(slope.c, speed.c)
```

```{r}
start_cities <- cities %>%
  select(lon, lat) %>%
  as.matrix

acc <- accCost(Conductance.c, start_cities) %>%
  mask(srtm)

acc %>%
  `/`(3600) %>%
  as.data.frame(xy = T, na.rm = T) %>%
ggplot(aes(x, y)) +
  geom_tile(aes(fill = layer)) +
  geom_polygon(data = centuriation, aes(long, lat, group = group), fill = 'lightgrey') +
  theme_void() +
  geom_point(data = places %>% filter(str_detect(type, 'settlement|villa')), aes(lon, lat), size = .5, alpha = .5) +
    geom_path(data = roads, aes(long, lat, group = group), alpha = .1) +
    geom_point(data = cities, aes(lon, lat, size = pop_est)) +
  scale_fill_viridis(direction = -1) +
  coord_quickmap() +
  theme(legend.position="none")
```

```{r}
library(tidygraph)
library(ggraph)

dist.mat <- costDistance(Conductance.c, start_cities) %>%
  `/`(3600) %>%
  as_tbl_graph(diagonal = F) %>%
  mutate(name = cities$Name, population = 1, attractiveness = 1, x = cities$lon, y = cities$lat, pop = cities$population, pop_est = cities$pop_est, coastal = cities$coastal) %E>%
  group_by(from) %>%
  rename(distance = weight) %>%
  filter(row_number(distance) <= 10) %>%
  ungroup


interact <- function(net){
  net %E>%
    mutate(interaction_strength = .N()$attractiveness[to] ^ alpha * exp(-beta * distance))  %N>%
    mutate(outflow = population / centrality_degree(weights = interaction_strength, mode = 'out', loops = F)) %E>%
    mutate(flow = .N()$outflow[from] * interaction_strength) %N>%
    mutate(inflow = centrality_degree(weights = flow, mode = 'in', loops = F)) %>%
    mutate(attractiveness = attractiveness + .01 * (inflow * coast_effect * coastal + inflow  - attractiveness)) %>%
    mutate(population = 97 * attractiveness / sum(attractiveness))
}

alpha <- 1.05
beta <- .57
coast_effect <- .5

net <- dist.mat
for(i in 1:500){
  net <- interact(net)
}

net %E>%
 group_by(from) %>%
  filter(flow == max(flow), .N()$population[from] < .N()$population[to]) %N>%
  mutate(terminal = node_is_sink()) %>%
ggraph(layout = 'manual', node.positions = as_tibble(activate(dist.mat, nodes))) +
  geom_edge_link(alpha = .6) +
  geom_node_point(aes(size = population, color = terminal)) +
 # geom_node_label(aes(label = name)) +
  scale_size_area() +
  coord_quickmap() +
  theme_void()
```


```{r}
ggraph(net, layout = 'manual', node.positions = as_tibble(activate(dist.mat, nodes))) +
  geom_edge_link(alpha = .1) +
  geom_node_point(aes(size = pop_est / mean(pop_est))) +
  scale_size_area() +
  coord_quickmap() +
  theme_void()
```

```{r cache = T}
costAll <- function(fromCoords, cost){
  
  aC <- srtm
  for(i in 1:dim(fromCoords)[1]) {aC <- stack(aC, accCost(cost, fromCoords[i,]))}
  aC <- dropLayer(aC,1)
  alC <- calc(aC, which.min)
  return(alC)
}

#https://stat.ethz.ch/pipermail/r-sig-geo/2011-July/012208.html
  
cum.cost <- net %E>%
 group_by(from) %>%
  filter(flow == max(flow), .N()$population[from] < .N()$population[to]) %N>%
  filter(node_is_sink()) %>%
  as_tibble %>%
  select(x, y) %>%
  as.matrix %>%
  costAll(Conductance.c) %>%
  mask(srtm) 
```

```{r}
terminal <- net %E>%
 group_by(from) %>%
  filter(flow == max(flow), .N()$population[from] < .N()$population[to]) %>%
    ungroup %N>%
    mutate(terminal = group_components()) %>%
  as_tibble %>%
  select(terminal) %>%
    cbind(1:97, .)

test <- cities %>%
 select(lon, lat) %>%
  as.matrix %>%
  costAll(Conductance.c) %>%
  mask(srtm) %>%
  reclassify(terminal) %>%
  rasterToPolygons(dissolve = T) %>% 
  fortify

net %E>%
 group_by(from) %>%
  filter(flow == max(flow), .N()$population[from] < .N()$population[to]) %N>%
  mutate(terminal = node_is_sink()) %>%
ggraph(layout = 'manual', node.positions = as_tibble(activate(dist.mat, nodes))) +
  geom_path(data = test, aes(long, lat, group = group), alpha = .1) +
  geom_point(data = places %>% filter(str_detect(type, 'settlement|villa')), aes(lon, lat), size = .5, alpha = .2) +
  geom_edge_link(alpha = .6) +
  geom_node_point(aes(size = population, color = terminal)) +
 # geom_node_label(aes(label = name)) +
  scale_size_area() +
  coord_quickmap() +
  theme_void() +
  theme(legend.position="none")

```

